---
title: 'W6: Practicing Functions'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(gganimate)
library(gapminder)
```
Define a defensive function that calculates the Gross Domestic Product of a nation from the data available in the gapminder dataset. Using that function, calculate the GDP of Denmark in the following years: 1967, 1977, 1987, 1997, 2007 and 2017.
#OBS! 2017 does not exist in the dataset

```{r}
#Creating a function, which calculates the GDP per nation by multiplying the "GDP per capita" with "population":

calcGDP <- function(dat) {
  gdp <- dat$pop * dat$gdpPercap
  return(gdp)
}

calcGDP <- function(dat, year=NULL, country=NULL) {
 if(!is.null(year)) {
    dat <- dat[dat$year %in% year, ]
  }
  if (!is.null(country)) {
    dat <- dat[dat$country %in% country,]
  }
  gdp <- dat$pop * dat$gdpPercap

  new <- cbind(dat, gdp=gdp)
  return(new)
}

# Printing the GDP for Denmark in specific years:
calcGDP(gapminder, year=c(1967, 1977, 1987, 1997, 2007, 2017), country="Denmark")
```


Write a script that loops over each country in the gapminder dataset, tests whether the country starts with a ‘B’ , and print out whether the life expectancy is smaller than 50, between 50 and 70, or greater than 70.

```{r}
# I grep the countries, which starts with B and put them into an object called BCountries
unique(gapminder$country)
BCountries <- grep("^B", unique(gapminder$country), value=TRUE) 

```


```{r}
# The countries started with B listed
BCountries
```

```{r}
# I don't understand why this is necessary, but it doesn't work without (Help??):
gapminder <- as.data.frame(gapminder) 


# Creating a loop that checks the average Life Expectancy for the countries starting with B, whereafter it prints them 

for (iCountry in BCountries) {
  tmp <- mean(gapminder[gapminder$country == iCountry, "lifeExp"])
  if(tmp >= 70){
    print(paste("Average Life Expectancy in", iCountry, "is equal to or larger than")) }  
  else if (tmp>=50) {
   print(paste("Average Life Expectancy in", iCountry, "is between 50 and 70 years"))}
  else { print(paste("Average Life Expectancy in", iCountry, "is smaller than 50"))
  }}
```



