---
title: "Homework week 8"
author: "Anna Hald"
date: "2/11/2020"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## GOAL

1) adapt the webscraping example to scrape homicide data from FBI site and produce a meaningful report on how homicide trends evolve around US in relation to this urban unrest

I use data from following website:
https://ucr.fbi.gov/crime-in-the-u.s/2018/crime-in-the-u.s.-2018/topic-pages/tables/table-4 

# Solution
First, load a handful of classic R packages:

{rvest} for web-scraping
{dplyr} for data-wrangling
{tidyr} for data transformation
{stringr} for string manipulation
{janitor} for clean headers that your OCD will love you for


```{r libraries}
library(tidyverse)
library(rvest)
library(dplyr)
library(tidyr)
library(stringr)
library(janitor)
```

## Scrape the data

Next, learn how scrape the content of the website and extract the HTML table:
```{r}
url <- "https://ucr.fbi.gov/crime-in-the-u.s/2018/crime-in-the-u.s.-2018/topic-pages/tables/table-4"
# scrape the website
url_html <- read_html(url)

whole_table <- url_html %>% 
 html_nodes("table") %>%
 html_table(fill = TRUE)  #str(whole_table) turns out to be a list

```

I have found guidance from following tutorial: https://www.r-bloggers.com/2020/04/scrape-html-table-using-rvest/

```{r}
# Changing the table from a list into a dataframe:
table <- whole_table[[1]]

glimpse(table) #shows that it reads the table as characters instead of numbers

# Cleaning the table
table <- table[-1,] #removes first row after the coloumn names as it is not useful
table <- table %>% clean_names() #clean the names of the coloumns (a part of the janitor package)
names(table)
```

Now I have a useful table. I wish to look at the development of homicides from 2017 to 2018 around US. The FBI table already has a row, which shows the percent change from 2017 to 2018 in each state, and therefore I filter that out in the following chunk. Thereafter, I choose the relevant coloumns, arrange them and create a barplot to show the difference in development according to states.

```{r}
table %>% 
	filter(year == "Percent change") %>% 
	mutate(murder_and_nonnegligent_manslaughter = parse_number(murder_and_nonnegligent_manslaughter)) %>%  #this mutates from characters to numbers 
	select(area,murder_and_nonnegligent_manslaughter) %>% 
	arrange(desc(murder_and_nonnegligent_manslaughter)) %>%
  mutate(area = fct_reorder(area, murder_and_nonnegligent_manslaughter)) %>%
	head(35) %>% # I only take the head 35, because otherwise it gets crammed...
	ggplot() + geom_bar(aes(y = area, x = murder_and_nonnegligent_manslaughter), stat = "identity", fill = "yellow") +
labs(title = "Development of homicides from 2017 to 2018 in US States",
       caption = "Data Source: FBI:UCR", x = "Percent change in number of murders from 2017 to 2018", y = "Area")
```

