---
title: "Final Project"
author: "Anna Hald"
date: "3/11/2020"
output: html_document
			
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# GOAL: Analyze data from Boligstyreslsen's lists of 'Ghettoes' and 'Vulnerable areas' and visualize it in plots
Link to data: https://www.trafikstyrelsen.dk/da/Bolig/Udsatte-boligomraader/Udsatte-omraader-og-ghettoomraader#omraadeafgraensninger 

## PACKAGES
First, I load the necessary packages:

```{r packages}
library(tidyverse)
library(dplyr)
library(tidyr)
library(ggplot2)
library(knitr)
```

## DATA 
I read in the 3 cleaned csv-files (cleaned in Open Refine on beforehand). One for each list.
The last part colClasses = "character" is in order to read the data as characters and not as numbers, because otherwise the commas in the files are lost.

```{r tables}
ghettoer <- read.csv(file = "data/Ghettoer_clean.csv", colClasses = "character")
hårdeghettoer <- read.csv(file = "data/Hårde_clean.csv", colClasses = "character")
udsatte <- read.csv(file = "data/Udsatte_clean.csv", colClasses = "character")
```

I wish to test, if the values in the dataframes are read as numbers, even though I specified it as characters. And they are!
```{r}
ghettoer %>% 
	ggplot() + geom_bar(aes(x = Område, y = Etnicitet), stat = "identity", fill = "yellow")
```

## DATA WRANGLING 1
I wish to have a dataframe with all the areas ('vulnerable' & 'ghettoes') in which it is possible to see how the area is categorized. 
Therefore, I merge the dataframes I add to the dataframe 'udsatte' (vulnerable) the coloumn 'Etnicitet' from the dataframe 'ghetto'. 

```{r}
etnicitet <- ghettoer %>% 
	select(Område, Etnicitet)
samlet <- merge(x=udsatte, y=etnicitet, by = "Område", all.x = TRUE)
```

Now that I have a way to differentiate the ghettoes, I can use the mutate in combination with the if-else function to create a coloumn that adds the value 'ghetto', if the ethnicity is > 50, and 'vulnerable area' if it is under.

```{r}
samletkategoriseret <- mutate(samlet, Ghetto = ifelse(Etnicitet > 50, "''Ghetto''", "''Vulnerable Area''"))
```

It does not count in the NA's, so instead I add the value "Vulnerable Area" in this way:
```{r}
samletkategoriseret$Ghetto[is.na(samletkategoriseret$Ghetto)] <- "''Vulnerable Area''"
```

## I wish to further differentiate the ghettoes, so it shows which are "hard ghettoes"
When I try to merge the tables (the last line, 77), only one of the hard ghettoes appear. I cannot figure out why? Therefore, I do not use the new dataframes in the following plots... 
```{r}
HårdeGhettoer <- hårdeghettoer$Område
allareas <- samletkategoriseret %>%
	mutate(samletkategoriseret, Categorization = ifelse(samletkategoriseret$Område %in% HårdeGhettoer, "Hård Ghetto", "Ikke"))
	
hårdeghettoer_2 <- hårdeghettoer %>% 
	mutate(Kategoriseret = "Hård_Ghetto")

HÅRD <- hårdeghettoer_2 %>% 
	select(Område, Kategoriseret)

allcategorized <- merge(x=samletkategoriseret, y=HÅRD, by = "Område", all.x = TRUE)
```

## PLOTTING 1
I install a package with extra color pallettes, I can choose from, when plotting.
```{r}
# install.packages("RColorBrewer")
library(RColorBrewer)
display.brewer.all()
```

Now I can start plotting. I wish to have a plot that shows the crime rates for each area and which clearly shows, if the area is defined as a ghetto or not. I do this by using an argument that picks color according to the value of the 'Ghetto' coloumn.
Furthermore, I add ad intercepting line to mark where the threshold value is for the areas to meet the criteria.
I have used the following tutorial to receive help on customizing labels on the plot:
http://www.sthda.com/english/wiki/ggplot2-axis-ticks-a-guide-to-customize-tick-marks-and-labels

```{r, echo=FALSE}
p <- samletkategoriseret %>% 
	mutate(Område = fct_reorder(Område, Dømte)) %>%
	ggplot(aes(x = Område, y = Dømte, fill = Ghetto)) + # 'FILL = ghetto i stedet for 'color' udfylder barerne i stedet for kun at optegne omrids
  geom_bar(stat = "identity") + scale_fill_brewer(palette = "Pastel2") + #Colors from the scale_fill_brewer package
	coord_flip() + #FLIPS THE PLOT
	theme(axis.text.x = element_text(size=5, angle=45), axis.text.y = element_text(size=7, angle=0)) + #Changes the angle on the numbers
	geom_hline(yintercept=24, linetype="dashed", color = "coral") +
	labs(title="CRIME RATES (based on data from 2017-2018)",
         x="Residential Areas on the ''Vulnerable Areas'' list ",
         y="Percentage of Sentenced Residents")+
	guides(fill=guide_legend(title="Categorization of the Area"))
p + annotate("text", x=5, y=31, label="Areas ≥ 2.03 %", size=4, color="coral") + annotate("text", x=3, y=31, label="meet the criteria", size=4, color="coral")
```

## DATA WRANGLING 2
I wish to do an analysis to see how many areas that meet each of the 4 criteria.

First, I need to change the commas in the dataframes, so R can read it properly.
```{r}
samletkategoriseret$Dømte <- gsub(",",".", c(samletkategoriseret$Dømte))
samletkategoriseret$Uddannelse <- gsub(",",".", c(samletkategoriseret$Uddannelse))
samletkategoriseret$Indkomst <- gsub(",",".", c(samletkategoriseret$Indkomst))
samletkategoriseret$Etnicitet <- gsub(",",".", c(samletkategoriseret$Etnicitet))
samletkategoriseret$Ledighed <- gsub(",",".", c(samletkategoriseret$Ledighed))
```

Second, I create an object for each criteria using the length argument to summarize the number of areas that meet each criteria.
```{r}
n_dømte <- length(samletkategoriseret[samletkategoriseret$Dømte >=2.03, "Dømte"])
n_ledighed <- length(samletkategoriseret[samletkategoriseret$Ledighed > 40, "Ledighed"])
n_uddannelse <- length(samletkategoriseret[samletkategoriseret$Uddannelse > 60, "Uddannelse"])
n_indkomst <- length(samletkategoriseret[samletkategoriseret$Indkomst < 55, "Indkomst"])
```

I create a new dataframe that contains a coloumn with these lengths and a coloumn with the names of the criterias
```{r}
n_alle <- c(n_dømte, n_ledighed, n_uddannelse, n_indkomst)
kriterier <- c("Crime Rate","Employment","Education","Income")

criteriacount <- data.frame(n_alle, kriterier)
```

## PLOTTING 2
Now, it is easy to plot.
```{r, echo=FALSE}
criteriacount %>% 
	ggplot(aes(x=kriterier,y=n_alle, fill = kriterier)) + 
  geom_bar(stat = "identity") + theme_minimal() + scale_fill_brewer(palette = "Pastel2") +
	geom_text(aes(label=n_alle), vjust=1.6, color="white", size=3.5) +
	labs(title="NUMBER OF AREAS MEETING EACH CRITERIA",
         x="Criteria",
         y="Number of Areas")+
	guides(fill=guide_legend(title="Criteria"))
```

